---
title: "R is a functional language"
author: "Marco Pascucci"
date: "25/10/2018"
header-includes: \usepackage{bbm}
output:
  beamer_presentation: 
    fig_height: 4
    fig_width: 7
  ioslides_presentation: default
  slidy_presentation: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(size = 'small')
knitr::opts_chunk$set(result = FALSE)



def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

library(tidyverse)
library(magrittr)

kable <- function(t) {
  kableExtra::kable(t , format = "markdown")
}
```

## list of function syntax
```{r result=TRUE}
summary <- function(x) {
  funs <- c(mean, median, sd, mad, IQR)
  lapply(funs, function(f) f(x, na.rm = TRUE) )
}

y <- summary(1:10)
str(y)
```

## Closures
functions that write functions
```{r}
power <- function(exponent) {
  function(x) {
    x ^ exponent
  }
}

square <- power(2)
square(2)
square
```

## Mutable states
A closure remembers it's creation environment (namespace and values). 
The double arrow operator allows to work on symbols of the creation environment, instead of locals.
```{r result=TRUE, size='tiny'}
func_factory = function() {
  i <- 0
  # returns this function
  function () {
    i <<- i+1
    i
  }
}

# function created, i is initialised to zero
f <- func_factory()

# i is incremented at each call.
f()
f()
```

## note on names hiding
Works only in the forward direction.
```{r size='tiny'}
# x is acessible in the function environment
f <- function() { x }
x<-3
f()

# x<-5 is hidden from the outer environment
f <- function() { x<-5; x }
x<-3
f()

fun_factory <- function() {
  function () { i <<- i+1; i }
}
# now we have to define i, because it is not defined in the closure body
i<-0
f=fun_factory()
i <--6 # by the way, values are taken at function's call, not creation
f() 
f()
```


## import list of functions in the local namespace
4 methods exist
```{r size='tiny'}
simple_tag <- function(tag) {
  force(tag)
  function(...) {
    paste0("<", tag, ">", paste0(...), "</", tag, ">")
  }
}
# could have used setNames()
tags <- c(p="p", b="b", i="i")
#html is a list of functions
html <- lapply(tags, simple_tag)
# INSTEAD OF
html$p("This is ", html$b("bold"), " text.")

## One can use any of the following
with(html, p("This is ", b("bold"), " text."))

#attach(html); p("This is ", b("bold"), " text."); detach(html)

#list2env(html, environment())
#  p("This is ", b("bold"), " text.")
#rm(list = names(html), envir = environment())
```


# FORMULA
## Exercice
expliquer le code suivant:
```{r result=FALSE, eval=FALSE}
library(purrr)
iris %>% group_by(Species) %>%
nest(.key = Data) %>%
mutate(Model = purrr::map(Data, ~ lm(data = .,
  Sepal.Length ~ Petal.Length))) %>%
mutate(Summary = purrr::map(Model, summary)) %>%
mutate(`R squared` = purrr::map_dbl(Summary, ~ .$r.squared))
```

# comme les "fonctions anonimes"...
```{r}
increment <- function(x) {
  x+1
}

v = 0:9

v %>% increment
v %>% (function(x) {x+1})
```

# Formula et l'operateur ~
Il existe une autre syntaxe, la `formula`, qui décrit une relation entre des variables, mais à difference d'une fonction, elle est seulement formelle.
```{r}
y ~ x1 + x2
```
cette `formula` veut seulement dire `y` depend de `x1` et `x2`.

Certaines fonctions, comme par exemple `map` peut prendre des `formulae` aussi bien que des fonctions en argument.
```{r}
before <- 0:9
map(before, ~.x+1) %>% cbind(before, after=.)
```


# coercion

explicit coercion (casting)
```{r}
as.character(1)
as.logical(1)
```

implicit coercion happens, for example, when data of different class are inserted in the same vector.
```{r}
c(1, TRUE, "a")
```
in this example all becomes character.

Imprecise coercion results in `NA`
```{r}
as.numeric('f')
```

# Factors
```{r}
unordered <-factor(c("one", "two", "three"))
ordered <- factor(c("one", "two", "three"),
                  c("one", "two", "three"))

str(unordered)
str(ordered)


```

factor are assigned in alphabetical order